Visual Diagnostics for Constrained Optimisation with Application to Guided Tours

### 2021-09-28 paper/zhang-cook-laa-langrene-menendez.tex
misc fix as per issue 18

The type of optimisation problem considered in this paper is constrained
optimization \citep{bertsekas2014constrained}, assuming it is not
possible to find a solution to the problem in the way of a closed-form.
That is, the problem consists in finding the minimum or maximum of a
function \(f \in L^p\) in the constrained space \(\mathcal{A}\).

@@

As a general visualisation method, tour produces animations of
high-dimensional data via rotations of low-dimensional planes. There are
different versions depending on how the high-dimensional space is
investigated: grand tour \citep{cook2008grand} selects the planes
randomly to provide a general overview; manual tour
\citep{cook1997manual} gradually phases in and out one variable to
understand the contribution of that variable in the projection. Guided
tour, the main interest of this paper, chooses the planes with the aid
of projection pursuit, to gradually reveal the most interesting
projection. Given a random start, projection pursuit iteratively finds
bases with higher index values and the guided tour constructs a geodesic
interpolation between these planes to form a tour path. Figure
\ref{fig:tour-path} shows a sketch of the tour path where the blue
frames are produced by the projection pursuit optimisation and the white
frames are interpolations between the blue frames. Mathematical details
of the geodesic interpolation can be found in
\citet{buja2005computational}. The aforementioned tour method has been
implemented in the R package \CRANpkg{tourr} \citep{tourr}.

@@

{\centering \includegraphics[width=0.7\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/tour-path}

@@

\caption[An illustration for demonstrating the frames in a tour path]{An illustration for demonstrating the frames in a tour path. Each square (frame) represents the projected data with a corresponding basis. Blue frames are returned by the projection pursuit optimisation and white frames are constructed between two blue frames by geodesic interpolation.}\label{fig:tour-path}
\end{figure}
\end{Schunk}

@@

Three main pieces of information are recorded during the projection
pursuit optimisation: 1) projection bases \(\mathbf{A}\), 2) index
values \(I\), and 3) state \(S\). For CRS and SA, possible states
include \texttt{random\_search}, \texttt{new\_basis}, and
\texttt{interpolation}. Pseudo-derivative (PD) has a wider variety of
states including \texttt{new\_basis}, \texttt{direction\_search},
\texttt{best\_direction\_search}, \texttt{best\_line\_search}, and
\texttt{interpolation}. Multiple iterators index the information
collected at different levels: \(t\) is a unique identifier prescribing
the natural ordering of each observation; \(j\) and \(l\) are the
counter of the outer and inner loop respectively. Other parameters of
interest recorded include \texttt{method} that tags the name of the
optimiser, and \texttt{alpha} that indicates the sampling neighbourhood
size for searching observations. A matrix notation of the data structure
is presented in equation \ref{eq:data-structure}.

\begin{equation}
\renewcommand\arraystretch{2}  % default value: 1.0
\begin{pmatrix}
t & \mathbf{A} & I & S & j &  l  & V_{1} & V_{2}\\
\hline
1 & \mathbf{A}_1 & I_1 & S_1 & 1 & 1 & V_{11} & V_{12}\\
\hline
2 & \mathbf{A}_2 & I_2 & S_2 & 2 & 1  & V_{21}  & V_{22}\\
\vdots & \vdots &\vdots &\vdots  &\vdots & \vdots &\vdots  &\vdots\\
\vdots & \vdots & \vdots &\vdots & 2 & l_2 & \vdots  & \vdots\\
\hline
\vdots &\vdots & \vdots &\vdots & 2  & 1& \vdots & \vdots\\
\vdots &\vdots &\vdots &\vdots &\vdots & \vdots & \vdots  &\vdots \\
\vdots &\vdots &\vdots &\vdots & 2 & k_2 &\vdots  & \vdots\\
\hline
\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
\hline
\vdots & \vdots & \vdots &\vdots  & J &  1 & \vdots & \vdots \\
\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
T & \mathbf{A}_T & I_T &S_T  & J &  l_{J} & V_{T1}& V_{T2}\\
\hline
\vdots &\vdots & \vdots &\vdots & J  & 1& \vdots & \vdots\\
\vdots &\vdots &\vdots &\vdots &\vdots & \vdots & \vdots  &\vdots \\
\vdots &\vdots &\vdots &\vdots & J & k_J &\vdots  & \vdots\\
\hline
\vdots& \vdots & \vdots & \vdots & J+1 & 1 & \vdots& \vdots\\
\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
T^\prime & \mathbf{A}_{T^\prime} & I_{T^\prime} &S_{T^\prime}  & J+1 &  l_{J+1} & V_{T^\prime 1}& V_{T^\prime 2}\\
\end{pmatrix}
= 
\begin{pmatrix}
\text{column name} \\
\hline
\text{search (start basis)} \\
\hline
\text{search} \\
\vdots \\
\text{search (accepted basis)} \\
\hline
\text{interpolate} \\
\vdots \\
\text{interpolate} \\
\hline
\vdots \\
\hline
\text{search} \\
\vdots \\
\text{search (final basis)} \\
\hline
\text{interpolate} \\
\vdots \\
\text{interpolate} \\
\hline
\text{search (no output)} \\
\vdots \\
\text{search (no output)} \\
\end{pmatrix}
\label{eq:data-structure}
\end{equation}

@@

Figure \ref{fig:toy-search} shows an example of the search plot for CRS
(left) and SA (right). Both optimisers quickly find better bases in the
first few iterations and then take longer to find one in the later
iterations. The anchor bases, the ones found with the highest index
value in each iteration, always have an increased index value in the
optimiser CRS while this is not the case for SA. This feature gives CRS
an advantage in this simple example to quickly find the optimum.

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/toy-search}

@@

\caption[A comparison of the searches by two optimisers]{A comparison of the searches by two optimisers: CRS (left) and SA (right) on a 2D projection problem of a six-variable dataset, \code{boa6} using the holes index. Both optimisers reach the final basis with a similar index value while it takes SA longer to find the final basis. In the earlier iterations, optimisers quickly find a better basis to proceed while in the later iterations, most sampled bases fail to make an improvement on the index value and a boxplot is used to summarise the distribution of the index values. There is no better basis found in the last iteration, 9 (left) and 15 (right), before reaching the maximum number of try and hence it is colored grey. The color scale is from the customised botanical palette in the ferrn package.}\label{fig:toy-search}
\end{figure}
\end{Schunk}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/toy-interp-1}

@@

\caption[An inspection of the index values as the optimisation progress for two optimisers]{An inspection of the index values as the optimisation progress for two optimisers: CRS (left) and SA (right). The holes index is optimised for a 2D projection problem on the six-variable dataset \code{boa6}. Lines indicate the interpolation and dots indicate new target bases generated by the optimisers. Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index values before reaching the target bases in time 76-130.}\label{fig:toy-interp}
\end{figure}
\end{Schunk}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/toy-pca}

@@

\caption[Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \code{boa5} using holes index]{Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \code{boa5} using holes index. The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA. The black star represents the theoretical best basis the optimisers aiming to find. All the bases in PD have been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases.}\label{fig:toy-pca}
\end{figure}
\end{Schunk}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/toy-pca-animated-1}

@@

\caption[Six frames selected from the animated version of the previous plot]{Six frames selected from the animated version of the previous plot. With animation, the progression of the search paths from start to finish is better identified. CRS (green) finishes the optimisation quicker than PD (brown) since there is no further movement for CRS in the sixth frame. The full video of the animation can be found in the html version of the paper.}\label{fig:toy-pca-animated}
\end{figure}
\end{Schunk}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/toy-tour-1}

@@

\caption{Six frames selected from rotating the high-dimensional basis space, along with the same two search paths from Figure \ref{fig:toy-pca} and \ref{fig:toy-pca-animated}. The basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package geozoo. The full animation can be seen in the html version of the paper.}\label{fig:toy-tour}
\end{figure}
\end{Schunk}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/toy-torus-1}

@@

\caption[Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green)]{Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green). The projection problem is a 2D projection with three variables using the holes index. The grey points are randomly generated 2D projection bases in the space and it can be observed that these points form a torus. The full video of the animation can be found in the html version of the paper.}\label{fig:toy-torus}
\end{figure}
\end{Schunk}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/interruption-1}

@@

\caption[Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data using holes index, optimised by CRS]{Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data using holes index, optimised by CRS. On the left panel, basis with a higher index value is found during the interpolation but not used. On the right panel, the interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with a higher index value, as shown on the right panel.}\label{fig:interruption}
\end{figure}
\end{Schunk}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/polish-1}

@@

\caption{Comparison of the projected data before and after using polishing for a 2D projection problem on \code{boa6} data using holes index. The top row shows the initial projected data and the final views after CRS and polish search and the second row traces the index value. The clustering structure in the data is detected by CRS (top middle panel) but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertical axis. Note that the parameter \code{max.tries} is set to 400 in this experiment for CRS to do its best. }\label{fig:polish}
\end{figure}
\end{Schunk}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/noisy-better-geo}

@@

\caption[Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \code{boa5}]{Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \code{boa5}. Both CRS and SA succeed in the optimisation, PD fails to optimise this non-smooth index. Further, SA takes  much longer than CRS to finish the optimisation, but finishes off closer to the theoretical best.}\label{fig:noisy-better-geo}
\end{figure}
\end{Schunk}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/kol-result}

@@

\caption[Comparing 20 search paths in the PCA-projected basis space faceted by two optimisers]{Comparing 20 search paths in the PCA-projected basis space faceted by two optimisers: CRS and SA, and two search sizes: 0.5 and 0.7. The optimisation is on the 1D projection index, $I^{nk}(n)$, for \code{boa6} data, where a local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*).}\label{fig:kol-result}
\end{figure}
\end{Schunk}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/flip-sign}

@@

\caption[Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the target basis]{Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the target basis. Optimisation is on the 1D projection index, $I^{nk}(n)$, for boa6 data using CRS with seed 2463. The dots represent the target basis in each iteration and the path shows the interpolation. On the left panel, one target basis is generated with an opposite orientation to the current basis (hence appear on the other side of the basis space) and the interpolator crosses the origin to perform the interpolation. The right panel shows the same interpolation after implementing an orientation check and the undesirable interpolation disappears.}\label{fig:flip-sign}
\end{figure}
\end{Schunk}

@@

  \begin{itemize}
  \tightlist
  \item
    \texttt{explore\_trace\_search()} produces summary plots in Figure
    \ref{fig:toy-search}.
  \item
    \texttt{explore\_trace\_interp()} produces trace plots for the
    interpolation points in Figure \ref{fig:toy-interp}.
  \item
    \texttt{explore\_space\_pca()} produces the PCA plot of projection
    bases on the reduced space in Figure \ref{fig:toy-pca}. The animated
    version in Figure \ref{fig:toy-pca-animated} can be turned on via
    the argument \texttt{animate\ =\ TRUE}.
  \item
    \texttt{explore\_space\_pca()} produces the PCA plot of projection
    bases on the reduced space. Figure \ref{fig:toy-pca} includes the
    additional details of anchor and search bases, which can be turned
    on by the argument \texttt{details\ =\ TRUE}. The animated version
    in Figure \ref{fig:toy-pca-animated} is produced with argument
    \texttt{animate\ =\ TRUE}.
  \item
    \texttt{explore\_space\_tour()} produces animated tour view on the
    full space of the projection bases in Figure \ref{fig:toy-tour}.
  \end{itemize}
\item
  \texttt{get\_*()} extracts and manipulates certain components from the
  existing data object.

@@

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}

Three algorithms (creeping random search, simulated annealing, and
pseudo-derivative) used in projection pursuit guided tour optimisation.

\begin{algorithm}
\SetAlgoLined
  \SetKwInOut{input}{input}
  \SetKwInOut{output}{output}
    \input{$f(.)$, $\alpha_1$, $l_{\max}$, $\text{cooling}$} 
    \output{$\mathbf{A}_{l}$}
    Generate random start $\mathbf{A}_1$ and set $\mathbf{A}_{\text{cur}} \coloneqq \mathbf{A}_1$, $I_{\text{cur}} = f(\mathbf{A}_{\text{cur}})$, $j = 1$\;
  \Repeat{$\mathbf{A}_l$ is too close to $\mathbf{A}_{\text{cur}}$ in terms of geodesic distance}{
   Set $l = 1$\;
  \Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$}{
    Generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
    Compute $I_{l}  = f(\mathbf{A}_{l})$\;
    Update $l = l + 1$\;
  }
  Update $\alpha_{j+1} = \alpha_j * \text{cooling}$\;
  Construct the geodesic interpolation between $\mathbf{A}_{\text{cur}}$ and $\mathbf{A}_l$\; 
  Update $\mathbf{A}_{\text{cur}} = \mathbf{A}_l$ and $j = j + 1$\;
}
  \caption{Creeping random search (CRS)}
  \label{random-search}
\end{algorithm}

\begin{algorithm}
\SetAlgoLined
\Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$ or $P > U$}{
    Generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
    Compute $I_{l}  = f(\mathbf{A}_{l})$, $T(l) = \frac{T_0}{\log(l + 1)}$ and $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
    Draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
    Update $l = l + 1$\;
  }
  \caption{Simulated annealing (SA)}
  \label{simulated-annealing}
\end{algorithm}

\begin{algorithm}
\SetAlgoLined
\Repeat{$l > l_{\max}$ or $p_{\text{diff}} > 0.001$}{
  Generate $n$ random directions $\mathbf{A}_{\text{rand}}$ \;
  Compute $2n$ candidate bases deviate from $\mathbf{A}_{\text{cur}}$ by an angle of $\delta$ while ensuring orthogonality\;
  Compute the corresponding index value for each candidate bases\;
  Determine the search direction as from $\mathbf{A}_{\text{cur}}$ to the candidate bases with the largest index value\;
  Determine the step size via optimising the index value on the search direction over a 90 degree window\;
  Find the optimum $\mathbf{A}_{**}$ and compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
  Update $l = l + 1$\;
}
\caption{Pseudo-derivative (PD)}
\label{search-geodesic}
\end{algorithm}

>>>>

The type of optimisation problem considered in this paper is constrained
optimization \citep{bertsekas2014constrained}, assuming it is not
possible to find a solution to the problem in the way of a closed-form.
That is, the problem consists in finding the minimum or maximum of a
function \(f \in L^p\) in the constrained space \(\mathcal{A}\), where
\(L^p\) defines the vector space of function \(f\), whose \(p\)th power
are integrable.

@@

As a general visualisation method, tour produces animations of
high-dimensional data via rotations of low-dimensional planes. There are
different versions depending on how the high-dimensional space is
investigated: grand tour \citep{cook2008grand} selects the planes
randomly to provide a general overview; manual tour
\citep{cook1997manual} gradually phases in and out one variable to
understand the contribution of that variable in the projection. Guided
tour, the main interest of this paper, chooses the planes with the aid
of projection pursuit, to gradually reveal the most interesting
projection. Given a random start, projection pursuit iteratively finds
bases with higher index values and the guided tour constructs a geodesic
interpolation between these planes to form a tour path. Figure
@ref(fig:tour-path) shows a sketch of the tour path where the blue
frames are produced by the projection pursuit optimisation and the white
frames are interpolations between the blue frames. Mathematical details
of the geodesic interpolation can be found in
\citet{buja2005computational}. The aforementioned tour method has been
implemented in the R package \CRANpkg{tourr} \citep{tourr}.

@@

{\centering \includegraphics[width=0.7\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/tour-path}

@@

\caption[An illustration for demonstrating the frames in a tour path]{An illustration for demonstrating the frames in a tour path. Each square (frame) represents the projected data with a corresponding basis. Blue frames are returned by the projection pursuit optimisation and white frames are constructed between two blue frames by geodesic interpolation.}(\#fig:tour-path)
\end{figure}
\end{Schunk}

@@

Three main pieces of information are recorded during the projection
pursuit optimisation: 1) projection bases \(\mathbf{A}\), 2) index
values \(I\), and 3) state \(S\). For CRS and SA, possible states
include \texttt{random\_search}, \texttt{new\_basis}, and
\texttt{interpolation}. Pseudo-derivative (PD) has a wider variety of
states including \texttt{new\_basis}, \texttt{direction\_search},
\texttt{best\_direction\_search}, \texttt{best\_line\_search}, and
\texttt{interpolation}. Multiple iterators index the information
collected at different levels: \(t\) is a unique identifier prescribing
the natural ordering of each observation; \(j\) and \(l\) are the
counter of the outer and inner loop respectively. Other parameters of
interest recorded, \(V\), include \texttt{method} that tags the name of
the optimiser, and \texttt{alpha} that indicates the sampling
neighbourhood size for searching observations. A matrix notation of the
data structure is presented in the table below.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.23}}@{}}
\toprule
\(t\) & \(\mathbf{A}\) & \(I\) & \(S\) & \(j\) & \(l\) & \(V_{1}\) &
\(V_{2}\) & Notes \\
\midrule
\endhead
\(1\) & \(\mathbf{A}_1\) & \(I_1\) & \(S_1\) & 1 & 1 & \(V_{11}\) &
\(V_{12}\) & start basis \\
\(2\) & \(\mathbf{A}_2\) & \(I_2\) & \(S_2\) & 2 & 1 & \(V_{21}\) &
\(V_{22}\) & search \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{} & \ldots{} & \ldots{} \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & 2 & \(l_2\) & \ldots{} &
\ldots{} & search (accepted) \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & 2 & 1 & \ldots{} & \ldots{}
& interpolation \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{} & \ldots{} & \ldots{} \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & 2 & \(k_2\) & \ldots{} &
\ldots{} & interpolation \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{} & \ldots{} & \ldots{} \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \(J\) & 1 & \ldots{} &
\ldots{} & search \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{} & \ldots{} & \ldots{} \\
\(T\) & \(\mathbf{A}_T\) & \(I_T\) & \(S_T\) & \(J\) & \(l_J\) &
\(V_{T1}\) & \(V_{T2}\) & search (final) \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \(J\) & 1 & \ldots{} &
\ldots{} & interpolation \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{} & \ldots{} & \ldots{} \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \(J\) & \(k_J\) & \ldots{} &
\ldots{} & interpolation \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \(J+1\) & 1 & \ldots{} &
\ldots{} & search (last round) \\
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{} & \ldots{} & \ldots{} \\
\(T^{\prime}\) & \(\mathbf{A}_{T^{\prime}}\) & \(I_{T^{\prime}}\) &
\(S_{T^{\prime}}\) & \(J+1\) & \(l_{J+1}\) & \(V_{{T}^{\prime}1}\) &
\(V_{{T}^{\prime}2}\) & search (last round) \\
\bottomrule
\end{longtable}

@@

Figure @ref(fig:toy-search) shows an example of the search plot for CRS
(left) and SA (right). Both optimisers quickly find better bases in the
first few iterations and then take longer to find one in the later
iterations. The anchor bases, the ones found with the highest index
value in each iteration, always have an increased index value in the
optimiser CRS while this is not the case for SA. This feature gives CRS
an advantage in this simple example to quickly find the optimum.

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-search}

@@

\caption[A comparison of the searches by two optimisers]{A comparison of the searches by two optimisers: CRS (left) and SA (right) on a 2D projection problem of a six-variable dataset, \code{boa6} using the holes index. Both optimisers reach the final basis with a similar index value while it takes SA longer to find the final basis. In the earlier iterations, optimisers quickly find a better basis to proceed while in the later iterations, most sampled bases fail to make an improvement on the index value and a boxplot is used to summarise the distribution of the index values. There is no better basis found in the last iteration, 9 (left) and 15 (right), before reaching the maximum number of try and hence it is colored grey. The color scale is from the customised botanical palette in the ferrn package.}(\#fig:toy-search)
\end{figure}
\end{Schunk}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-interp-1}

@@

\caption[An inspection of the index values as the optimisation progress for two optimisers]{An inspection of the index values as the optimisation progress for two optimisers: CRS (left) and SA (right). The holes index is optimised for a 2D projection problem on the six-variable dataset \code{boa6}. Lines indicate the interpolation and dots indicate new target bases generated by the optimisers. Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index values before reaching the target bases in time 76-130.}(\#fig:toy-interp)
\end{figure}
\end{Schunk}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-pca}

@@

\caption[Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \code{boa5} using holes index]{Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \code{boa5} using holes index. The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA. The black star represents the theoretical best basis the optimisers aiming to find. All the bases in PD have been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases.}(\#fig:toy-pca)
\end{figure}
\end{Schunk}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-pca-animated-1}

@@

\caption[Six frames selected from the animated version of the previous plot]{Six frames selected from the animated version of the previous plot. With animation, the progression of the search paths from start to finish is better identified. CRS (green) finishes the optimisation quicker than PD (brown) since there is no further movement for CRS in the sixth frame. The full video of the animation can be found in the html version of the paper.}(\#fig:toy-pca-animated)
\end{figure}
\end{Schunk}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-tour-1}

@@

\caption[Six frames selected from rotating the high-dimensional basis space, along with the same two search paths from Figure \@ref(fig:toy-pca) and \@ref(fig:toy-pca-animated)]{Six frames selected from rotating the high-dimensional basis space, along with the same two search paths from Figure \@ref(fig:toy-pca) and \@ref(fig:toy-pca-animated). The basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package geozoo. The full animation can be seen in the html version of the paper.}(\#fig:toy-tour)
\end{figure}
\end{Schunk}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-torus-1}

@@

\caption[Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green)]{Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green). The projection problem is a 2D projection with three variables using the holes index. The grey points are randomly generated 2D projection bases in the space and it can be observed that these points form a torus. The full video of the animation can be found in the html version of the paper.}(\#fig:toy-torus)
\end{figure}
\end{Schunk}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/interruption-1}

@@

\caption[Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data using holes index, optimised by CRS]{Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data using holes index, optimised by CRS. On the left panel, basis with a higher index value is found during the interpolation but not used. On the right panel, the interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with a higher index value, as shown on the right panel.}(\#fig:interruption)
\end{figure}
\end{Schunk}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/polish-1}

@@

\caption{Comparison of the projected data before and after using polishing for a 2D projection problem on \code{boa6} data using holes index. The top row shows the initial projected data and the final views after CRS and polish search and the second row traces the index value. The clustering structure in the data is detected by CRS (top middle panel) but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertical axis. Note that the parameter \code{max.tries} is set to 400 in this experiment for CRS to do its best. }(\#fig:polish)
\end{figure}
\end{Schunk}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/noisy-better-geo}

@@

\caption[Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \code{boa5}]{Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \code{boa5}. Both CRS and SA succeed in the optimisation, PD fails to optimise this non-smooth index. Further, SA takes  much longer than CRS to finish the optimisation, but finishes off closer to the theoretical best.}(\#fig:noisy-better-geo)
\end{figure}
\end{Schunk}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/kol-result}

@@

\caption[Comparing 20 search paths in the PCA-projected basis space faceted by two optimisers]{Comparing 20 search paths in the PCA-projected basis space faceted by two optimisers: CRS and SA, and two search sizes: 0.5 and 0.7. The optimisation is on the 1D projection index, $I^{nk}(n)$, for \code{boa6} data, where a local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*).}(\#fig:kol-result)
\end{figure}
\end{Schunk}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/flip-sign}

@@

\caption[Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the target basis]{Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the target basis. Optimisation is on the 1D projection index, $I^{nk}(n)$, for boa6 data using CRS with seed 2463. The dots represent the target basis in each iteration and the path shows the interpolation. On the left panel, one target basis is generated with an opposite orientation to the current basis (hence appear on the other side of the basis space) and the interpolator crosses the origin to perform the interpolation. The right panel shows the same interpolation after implementing an orientation check and the undesirable interpolation disappears.}(\#fig:flip-sign)
\end{figure}
\end{Schunk}

@@

  \begin{itemize}
  \tightlist
  \item
    \texttt{explore\_trace\_search()} produces summary plots in Figure
    @ref(fig:toy-search).
  \item
    \texttt{explore\_trace\_interp()} produces trace plots for the
    interpolation points in Figure \ref{fig:toy-interp}.
  \item
    \texttt{explore\_space\_pca()} produces the PCA plot of projection
    bases on the reduced space in Figure \ref{fig:toy-pca}. The animated
    version in Figure \ref{fig:toy-pca-animated} can be turned on via
    the argument \texttt{animate\ =\ TRUE}.
  \item
    \texttt{explore\_space\_pca()} produces the PCA plot of projection
    bases on the reduced space. Figure \ref{fig:toy-pca} includes the
    additional details of anchor and search bases, which can be turned
    on by the argument \texttt{details\ =\ TRUE}. The animated version
    in Figure \ref{fig:toy-pca-animated} is produced with argument
    \texttt{animate\ =\ TRUE}.
  \item
    \texttt{explore\_space\_tour()} produces animated tour view on the
    full space of the projection bases in Figure \ref{fig:toy-tour}.
  \end{itemize}
\item
  \texttt{get\_*()} extracts and manipulates certain components from the
  existing data object.

@@

\begin{verbatim}
\end{verbatim}
